A North Sámi to South Sámi machine translation prototype

### 2014-02-09 paper/paper.tex
more

@@

\textbf{Bilingual dictionary}

@@

\subsubsection{Problems caused by this model: ex from two terms to one term to two terms}

@@

\subsubsection{Problems combining sma-nob and sme-nob dictionary}
This was done, it had to be manually edited because of nob-words with more than one meaning. 
Example: nob \textit{regne} means both 'to rain' and ‘to calculate’, 
which are two different verbs both in sme \textit{arvit, rehkenastit} and in sma \textit{abrodh, ryöknedh}

\subsubsection{Rules}

@@

@@

>>>>

Morphological analysis is done on the input using the Helsinki Finite-State
Toolkit \cite{linden2011}. For each surface form, a finite-state transducer
returns a set of the possible analyses, where an analysis is a combination
of lemma, and a sequence of tags which describe the morphological structure
of the surface form.

A Constraint Grammar-based disambiguator then selects the most appropriate 
analysis for each surface form according to the context, and assigns to each 
analysis a syntactic tag denoting its syntactic function (subject, object, 
main verb, \ldots).

@@

\subsubsection{Lexical transfer}

@@

\textbf{Problems caused by this model: ex from two terms to one term to two terms}

@@

\textbf{Problems combining sma-nob and sme-nob dictionary}
This was done, it had to be manually edited because of nob-words with more than one meaning. 
Example: nob \textit{regne} means both 'to rain' and ‘to calculate’, 
which are two different verbs both in sme \textit{arvit, rehkenastit} and in sma \textit{abrodh, ryöknedh}

\subsubsection{Lexical selection}

\subsubsection{Structural transfer}

@@

Generation is done using a finite-state transducer, also 
compiled using HFST. Each lexical unit which is output by
the postchunk module is looked up in the generation transducer
and the surface form is output.

@@

Linda Wiechetek, Divvun
